---
title: "Identifying Cross-Selling Opportunities"
author: "Shubhankar Gupta"
date: "March 8, 2016"
output: 
  html_document: 
    keep_md: yes
---
I have used a Market Basket Analysis in R for the required solution.

Reading the data into R.
```{r, cache=TRUE}
product = read.table("ProductCategoryData.txt", header = T)
user = read.table("UserProfileData.txt", sep = "\t", header = T)
click = read.table("ClickStream.txt", sep = "\t", header=T)
```

State/Region of Portugal have been recorded as numeric values. Therefore I have removed them for easier calculation. Also futher cleaning the data.
```{r, cache=TRUE}
click = click[,-(8:16)]
x = which(click$Region.State=="10")
b = which(click$Region.State=="11")
c = which(click$Region.State=="47")
d = which(click$Region.State=="")
click$Region.State = as.character(click$Region.State)
click = click[-c(x,b,c,d),]
click$Region.State = as.factor(click$Region.State)
click$Country = as.character(click$Country)
click$Country = as.factor(click$Country)
click$City = as.character(click$City)
click$City = as.factor(click$City)
```

We merge the data of product category and click stream by URL, to access the click on a particular product category.
```{r}
a1 = merge(click, product, by= "URL")
```

Now we merge the data of a1 and user profile by USER ID, to access the user who is clicking for the product category.
```{r}
a2 = merge(a1, user, by = "User.ID", all.x = T)
```

Market Basket Analysis is also called product association analysis.Association mining analysis mostly done based on an algorithm named "Apriori Algorithm".

```{r}
library(arules)
```

Now I converted the data frame into transaction class as per the requirements of **arules** package.
```{r, cache=TRUE}
a3 = a2[,-3]
data = a3[,(1:2)]
r = unique(data)
head(r)
spl = split(r$URL, r$User.ID)
trans = as(spl, "transactions")
df = data.frame(URL=levels(r$URL))
df$URL =as.character(df$URL)
level = merge(df,product, by="URL")
```

Using **support** for this transaction as 0.001 and **confidence** as 0.75 to get the association rules. As we will decrease confidence number of association rules will increase.
```{r}
rules = apriori(trans, parameter = list(supp=0.001, conf=0.75))
inspect(rules)
```

Where these numbers correspond to a product category.
```{r}
level
```

Here are some Plots of the Rules we Obtained from the analysis.
```{r}
library(arulesViz)
plot(rules)
plot(rules, method="graph", control=list(type="items"))
plot(rules, method="paracoord", control=list(reorder=TRUE))

```


